---
title: "Assignment #2 — Statistical analysis of thermal comfort research"
author: "Kaito Funada"
date: 9/27/2026
format:
  html:
    toc: true
    code-fold: show
    code-tools: true
    embed-resources: true
execute:
  warning: false
  message: false
---

<!--
HOW TO USE THIS FILE
1. Click "Render" in RStudio (or run: quarto render analysis.qmd).
   If it renders before you change anything, your setup works.
2. Answer the eight questions below. Replace every [WRITE ...] prompt.
3. Every number in your text must come from a code chunk, not from
   typing it in by hand. That is the whole point of Quarto.
4. Render one more time before you submit, from a fresh R session
   (Session > Restart R). If it only renders in your current session,
   it is not reproducible.
-->

```{r setup}
library(ggplot2)
library(dplyr)
library(tidyr)

theme_set(theme_minimal(base_size = 12))

load("data/arch241a2.rda")
d <- arch241a2
```

## About this dataset — read before you start

Thirty subjects each sat through seven conditions in a climate chamber in
Guangzhou, giving 30 × 7 = 210 rows.

| Variable | Meaning | Notes |
|---|---|---|
| `Top` | Operative temperature, °C | Equals air temperature here (MRT = Ta) |
| `RH` | Relative humidity, % | Two levels: 50 and 70 |
| `Sex` | F / M | 15 subjects each |
| `Subject` | Subject ID, 1–30 | Each ID appears 7 times|
| `TS` | Thermal sensation | ASHRAE scale, −4 cold … 0 neutral … +4 hot |
| `TC` | Thermal discomfort | 0 = very comfortable … 4 = very uncomfortable |
| `TA` | Thermal acceptability | Higher = more acceptable |
| `PMV` | Predicted mean vote | Calculated from the design conditions |
| `PPD` | Predicted % dissatisfied | Calculated from PMV |
| `Experiment` | Condition A–F, plus P | P is pre-conditioning, not a test condition |

```{r orient}
str(d)
```


## 1. Statistical summary of all variables

```{r q1}
#  Summarise every variable.
summary(d)

# Think: does a mean make sense for a factor? For a design variable like Top?
```

**Interpretation.** [WRITE 2–3 SENTENCES. What would you put in a paper?]
The dataset contains `r nrow(d)` observations from
`r n_distinct(d$Subject)` subjects across `r n_distinct(d$Experiment)`
experimental conditions. The sample includes
`r n_distinct(d$Subject[d$Sex == "F"])` female and
`r n_distinct(d$Subject[d$Sex == "M"])` male subjects. Operative temperature
ranges from `r min(d$Top)` to `r max(d$Top)` °C, and thermal sensation ranges
from `r min(d$TS)` to `r max(d$TS)`.

## 2. Histograms of TC and TS

```{r q2}
qplot(TC, data=arch241a2, geom="histogram", binwidth=0.25) +
  xlab("Thermal comfort (TC)") +
  ylab("Count") +
  theme_bw()

qplot(TS, data=arch241a2, geom="histogram", binwidth=0.25) +
  xlab("Thermal sensation (TS)") +
  ylab("Count") +
  theme_bw()
```

**Interpretation.** [WRITE 2–3 SENTENCES]
TS appears roughly symmetric, while
TC looks not symmetric distribution. 
This suggests that TS is closer to a
normal distribution than TC.


## 3. Boxplots of TS vs Sex, Subject and Experiment

```{r q3}
# [WRITE YOUR CODE] Three plots.
qplot(Sex, TS, data=arch241a2, geom="boxplot") +
  xlab("Sex") +
  ylab("Thermal sensation (TS)") +
  theme_bw()

qplot(Subject, TS, data=arch241a2, geom="boxplot") +
  xlab("Subject") +
  ylab("Thermal sensation (TS)") +
  theme_bw()

qplot(Experiment, TS, data=arch241a2, geom="boxplot") +
  xlab("Experiment") +
  ylab("Thermal sensation (TS)") +
  theme_bw()
# The Subject one is the least obvious and the most informative — look at it carefully.
```

**Interpretation.** [WRITE 2–3 SENTENCES]
TS difference is almost same between female and male.
Almost all TS scatters around 0 for all subject, but not sure if there is a
significant difference.
TS looks different among each experiment condition.

## 4. Correlation matrix

Variables: PMV, PPD, RH, TA, TC, TS, Top.

```{r q4}
# [WRITE YOUR CODE]
corrTab <- cor(
  arch241a2[,c("PMV","PPD","RH","TA","TC","Top","TS")],
  use="complete",
  method="pearson"
)

round(corrTab,2)

library('GGally')

ggcorr(
  arch241a2[,c("PMV","PPD","RH","TA","TC","Top","TS")],
  label=TRUE
)

# Top vs PMV
ggplot(
  distinct(arch241a2, Top, PMV),
  aes(x=Top, y=PMV)
) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) +
  theme_bw()

# PMV vs PPD
ggplot(
  distinct(arch241a2, PMV, PPD),
  aes(x=PMV, y=PPD)
) +
  geom_point() +
  geom_line() +
  theme_bw()


```

**Interpretation.** [WRITE 2–3 SENTENCES]
TS is strongly positively correlated with Top
(r = `r round(corrTab["TS","Top"],2)`) and PMV
(r = `r round(corrTab["TS","PMV"],2)`).
TC and TA are strongly negatively
correlated (r = `r round(corrTab["TC","TA"],2)`). 


Top and PMV are almost perfectly correlated because PMV was calculated 
from the experimental thermal conditions. 
In contrast, PMV and PPD are deterministically related but their
correlation is only `r round(corrTab["PMV","PPD"],2)` because their
relationship is nonlinear (U curve) rather than linear.

## 5. Normality of TC and TS

```{r q5}
# [WRITE YOUR CODE] A test AND a plot. Do not rely on the p-value alone.
shapiro.test(arch241a2$TC)
qqnorm(arch241a2$TC, main="TC")
qqline(arch241a2$TC)


shapiro.test(arch241a2$TS)
qqnorm(arch241a2$TS, main="TS")
qqline(arch241a2$TS)
```

**Interpretation.** [WRITE 2–3 SENTENCES. Are they normal? Does it matter, and for what?]
TC significantly deviates from a normal distribution
(p = `r signif(shapiro.test(arch241a2$TC)$p.value, 3)`), and its Q-Q plot also
shows deviations from the reference line. 
In contrast, normality is not
rejected for TS (p = `r round(shapiro.test(arch241a2$TS)$p.value, 2)`).
Non-normality does not matter, but it will matter if TC is used in statistical test that
assume normality. 

## 6. Does Sex significantly influence TS?

```{r q6}
# [WRITE YOUR CODE]
# Then answer this, in your text: Sex is a property of a SUBJECT, and you have
# 30 subjects but 210 rows. What is your real sample size, and what does using
# 210 do to your p-value?
# Try at least two approaches and report whether they agree.

# Approach1: using 210 observations
t.test(
  TS ~ Sex,
  alternative='two.sided',
  conf.level=.99,
  var.equal=FALSE,
  data=arch241a2
)

# Approach 2: using the mean TS of 30 subjects
t.test(
  TS ~ Sex,
  alternative='two.sided',
  conf.level=.99,
  var.equal=FALSE,
  data=summarise(
    group_by(arch241a2, Subject, Sex),
    TS=mean(TS)
  )
)
```

**Interpretation.** [WRITE 2–3 SENTENCES]
Sex is a characteristic of the subject, so the real
independent sample size is `r n_distinct(arch241a2$Subject)`, not
`r nrow(arch241a2)` observations.  
Treating all observations as independent could ignore the 
repeated measurements within each subject.

Both approaches show no statistically
significant difference in TS between F and M.


## 7. Linear regression: TS as a function of Top

```{r q7}
# [WRITE YOUR CODE] Fit the model, then TEST ITS VALIDITY.
# The four assumptions: linearity, constant variance, normal residuals,
# independent observations. Check all four and say which ones hold.


lmTSTop <- lm(TS ~ Top, data=arch241a2)
summary(lmTSTop)

# Diagnostic plots
par(mfrow=c(2,2))
plot(lmTSTop)
par(mfrow=c(1,1))

# Normality of residuals
shapiro.test(residuals(lmTSTop))

# Regression plot
LinMod <- ggplot(
  arch241a2,
  aes(x=Top, y=TS)
) +
  geom_point() +
  theme_bw() +
  xlab("Operative temperature, Top [°C]") +
  ylab("Thermal sensation, TS")

print(LinMod + geom_smooth(method="lm"))

```

**Interpretation.** [WRITE 2–3 SENTENCES. Report the slope with units, the R²,
and the neutral temperature — the Top at which predicted TS = 0. Compare it to
what PMV predicts.]

Linearity, constant-variance and normal residuals appears reasonably appropriate.
Independence is not satisfied because each subject was measured repeatedly 
across seven conditions.


The slope is `r round(coef(lmTSTop)["Top"],3)` with units for each 1 °C increase
in Top.
The regression model explained
`r round(summary(lmTSTop)$r.squared*100,1)`% of the variation in TS.

The predicted neutral temperature was
`r round(-coef(lmTSTop)["(Intercept)"]/coef(lmTSTop)["Top"],1)` °C.

In PMV, it is
`r round(-coef(lm(PMV ~ Top, data=arch241a2))["(Intercept)"] /
coef(lm(PMV ~ Top, data=arch241a2))["Top"],1)` °C.

## 8. Your own research question

**Question:** [WRITE YOUR QUESTION]
Does experimental condition have a significant influence on thermal sensation?

```{r q8}
# [WRITE YOUR CODE]

AnovaTSExperiment <- aov(TS ~ Experiment, data=arch241a2)
summary(AnovaTSExperiment)

by_Experiment <- group_by(arch241a2, Experiment)
summarise(by_Experiment,
          count = n(),
          mean(TS),
          sd(TS))
```

**Interpretation.** [WRITE 2–3 SENTENCES — the take-away message you would put
in a paper.]
Thermal sensation differed significantly among experimental
conditions. 
Mean TS ranged from
`r round(mean(arch241a2$TS[arch241a2$Experiment=="A"]),2)` in condition A to
`r round(mean(arch241a2$TS[arch241a2$Experiment=="F"]),2)` in condition F,
showing a clear shift from cooler to warmer thermal sensations across conditions.


## Session information

```{r session}
sessionInfo()
```
